Cauchy example

### 2020-05-03 src/s_ar1.Rmd
Up to end of fig. 8

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
```

@@

  temp <- f_replicate()
  r_star_vals[i] <- temp$r_star
  r_hat_vals[i] <- temp$r_hat
}
```

Plot results
```{r}
g2 <- tibble(accuracy=r_star_vals, iteration=seq(1, nreplicates, 1)) %>% 
  ggplot(aes(x=accuracy)) +
  geom_histogram() +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  ggtitle("B.") +
  theme(text = element_text(size=14, colour="black")) +
  xlim(1, NA)
```

Look at predictions of model
```{r}
g1 <- testing_data %>% 
  arrange(value) %>% 
  mutate(predictions=predict(object=gbmFit1, .)) %>% 
  ggplot(aes(x=value, y=predictions)) +
  geom_jitter(height = 0.3) +
  xlab("Value") +
  ylab("Chain classification") +
  theme(text = element_text(size=14, colour="black")) +
  coord_flip() +
  ggtitle("A.")
```

Generate R* distribution based on 1000 iterations for run where chain has lower variance and another where all four chains have same variance
```{r}
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) 
a_array <- array(dim=c(1000, 4, 1))
a_array[,,1] <- temp
unmixed <- r_star_dist(a_array)

full_data <- f_generate_lower_var_four(1, 0.3, 1, 1000)
a_array <- array(dim=c(1000, 4, 1))
a_array[,,1] <- temp
mixed <- r_star_dist(a_array)

mAR1 <- tibble(unmixed, mixed)
g3 <- 
  mAR1 %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series") +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))

# pdf("../output/ar1.pdf", width = 12, height = 6)
multiplot(g1, g2, g3, cols = 3)
# dev.off()
```

>>>>

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

@@

  temp <- f_replicate()
  r_star_vals[i] <- temp$r_star
  r_hat_vals[i] <- temp$r_hat
}

saveRDS(tibble(r_star=r_star_vals, r_hat=r_hat_vals), "../data/ar1_r_star_hat.rds")
```

Plot results
```{r}
g2 <- tibble(accuracy=r_star_vals, iteration=seq(1, nreplicates, 1)) %>% 
  ggplot(aes(x=accuracy)) +
  geom_histogram() +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  ggtitle("B.") +
  theme(text = element_text(size=14, colour="black")) +
  xlim(0.9, NA)

g2a <- tibble(accuracy=r_hat_vals, iteration=seq(1, nreplicates, 1)) %>% 
  ggplot(aes(x=accuracy)) +
  geom_histogram() +
  xlab(TeX("split-$\\hat{R}$")) +
  ylab("Count") +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black")) +
  xlim(1, NA)
```

Look at predictions of model
```{r}
# add in a column of noise since gbm requires >1 variables
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  select(-Var1) %>% 
  rename(chain=Var2) %>% 
  mutate(ran_val=rnorm(4000)) %>% 
  mutate(chain=as.factor(chain))

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)

rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

g1 <- testing_data %>% 
  arrange(value) %>% 
  mutate(predictions=predict(object=gbmFit1, .)) %>% 
  ggplot(aes(x=value, y=predictions)) +
  geom_jitter(height = 0.3) +
  xlab("Value") +
  ylab("Chain classification") +
  theme(text = element_text(size=14, colour="black")) +
  coord_flip() +
  ggtitle("A.")
```

Generate R* distribution based on 1000 iterations for run where chain has lower variance and another where all four chains have same variance
```{r}
temp <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) 
temp1 <- matrix(rnorm(4000), nrow = 1000)
a_array <- array(dim=c(1000, 4, 2))
a_array[,,1] <- temp
a_array[,,2] <- temp1
unmixed <- r_star_dist(a_array)

temp <- f_generate_lower_var_four(1, 0.3, 1, 1000)
temp1 <- matrix(rnorm(4000), nrow = 1000)
a_array <- array(dim=c(1000, 4, 2))
a_array[,,1] <- temp
a_array[,,2] <- temp1
mixed <- r_star_dist(a_array)

mAR1 <- tibble(unmixed, mixed)
g3 <- 
  mAR1 %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("un-mixed", "mixed")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("D.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))

pdf("../output/ar1.pdf", width = 12, height = 6)
multiplot(g1, g2a, g2, g3, cols = 2)
dev.off()
```

### 2020-05-03 src/s_cauchy.Rmd
Up to end of fig. 8

r_star_nom <- r_star_dist(x)
```

@@

r_star_alt <- r_star_dist(full_data)
```

Plot split-R* values for each parameterisation
```{r}
a_df <- tibble(nominal=r_star_nom,
               alt=r_star_alt)
g <-
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        legend.position = "none")
g
```

## Calculate KL divergence from true density by fitting kernel density estimator to points then integrating
nominal parameterisation
```{r}
x <- rstan::extract(fit_nom, permuted=F)

sample <- as.vector(x[, , i])
dens <- density(sample)
f_approx <- approxfun(dens$x, dens$y, rule=2)
integrate(
    function(x) f_approx(x) * log(f_approx(x) / dcauchy(x)),      -20, 20, subdivisions = 1000)

t <- seq(-20, 20, 0.01)
tibble(t,
  true=map_dbl(t, ~dcauchy(.)),
  approx=map_dbl(t, ~f_approx(.))) %>%
  melt(id.vars="t") %>%
  ggplot(aes(x=t, y=value, colour=as.factor(variable))) +
  geom_line()

f_kl_approx <- function(sample, limit=20){
  dens <- density(sample)
  f_approx <- approxfun(dens$x, dens$y, rule=2)
  kl <- integrate(function(x) f_approx(x) * log(f_approx(x) / dcauchy(x)),      -limit, limit, subdivisions=1000)$value
  if(kl < 0)
    return(0)
  else
    return(kl)
}

kl <- vector(length = 50)
for(i in seq_along(kl))
  kl[i] <- f_kl_approx(x[, , i], limit=20)
```

Density estimation fails for alternative parameterisation because of fat tails
```{r}
names(extract(fit_alt1))
x <- rstan::extract(fit_alt1, permuted=F, pars="x")
sample <- as.vector(x[, , 3])
dens <- density(sample, 0.001)
f_approx <- approxfun(dens$x, dens$y, rule=2)
integrate(function(x) f_approx(x), -10, 10)
integrate(
    function(x) f_approx(x) * log(f_approx(x) / dcauchy(x)),      -100, 100, subdivisions = 1000)

t <- seq(-20, 20, 0.01)
tibble(t,
  true=map_dbl(t, ~dcauchy(.)),
  approx=map_dbl(t, ~f_approx(.))) %>%
  melt(id.vars="t") %>%
  ggplot(aes(x=t, y=value, colour=as.factor(variable))) +
  geom_line()

f_kl_approx <- function(sample, limit=20){
  dens <- density(sample)
  f_approx <- approxfun(dens$x, dens$y, rule=2)
  kl <- integrate(function(x) f_approx(x) * log(f_approx(x) / dcauchy(x)),      -limit, limit, subdivisions=1000)$value
  if(kl < 0)
    return(0)
  else
    return(kl)
}

kl <- vector(length = 50)
for(i in seq_along(kl))
  kl[i] <- f_kl_approx(x[, , i], limit=20)
```

Look at quantiles instead using KS test statistic: problem with KS test statistic is that it is most sensitive to middle values: https://asaip.psu.edu/articles/beware-the-kolmogorov-smirnov-test/
```{r}
x <- rstan::extract(fit_alt1, permuted=F, pars="x")
sample <- as.vector(x[, , 1])
ks.test(sample, pcauchy)
ks <- vector(length = 50)
for(i in 1:50)
  ks[i] <- ks.test(x[, , i], pcauchy)[1]$statistic
```

Anderson-darling test -- by comparing with an independent sampler we get a gauge of the quality of MCMC
```{r}
x <- rstan::extract(fit, permuted=F, pars="x")

f_calculate_ad <- function(x){
  dims <- dim(x)
  nparams <- dims[3]
  ad <- vector(length = nparams)
  for(i in 1:nparams)
    ad[i] <- ad.test(x[, , i], pcauchy)[2]$p.value
  return(ad)
}

# keep sample size fixed by thinning
nmax <- 5000
iter <- seq(500, nmax, 50)
ad <- vector(length = length(iter))
for(i in seq_along(iter)){
  a_iter <- iter[i]
  ele <- round(seq(1, a_iter, length.out = 500))
  ad[i] <- mean(f_calculate_ad(x[ele, ,]))
}

e_df <- tibble(iter, ad, source="MCMC")

# compare with independent samples
# ps <- map_dbl(seq(1, 1000, 1), function(i) ad.test(rcauchy(1000), pcauchy)[2]$p.value)
# hist(ps)

x <- array(rcauchy(2 * nmax * 50), c(nmax, 4, 50))

iter <- seq(500, nmax, 50)
ad <- vector(length = length(iter))
for(i in seq_along(iter)){
  a_iter <- iter[i]
  ele <- round(seq(1, a_iter, length.out = 500))
  ad[i] <- mean(f_calculate_ad(x[ele, ,]))
}
f_df <- e_df %>% 
  bind_rows(tibble(iter, ad, source="independent"))
f_df %>% 
  ggplot(aes(x=iter, y=ad, colour=as.factor(source))) +
  geom_line()
```

Try R* with independent sampler
```{r}
nmax <- 100
x <- array(rcauchy(2 * nmax * 50), c(nmax, 4, 50))
r_star_inde <- r_star_dist(x)
hist(r_star_inde)
```

Possible to tell whether sample came from MCMC or independent?
```{r}
x_mcmc <- rstan::extract(fit, permuted=F, pars="x")
x_mcmc <- array(rcauchy(2000 * 50), c(2000, 50)) %>% 
  as.data.frame() %>% 
  mutate(source="mcmc")

f_distribution <- function(x_mcmc){
  x_inde <- array(rcauchy(2000 * 50), c(2000, 50))
  x_inde <- x_inde %>% 
  as.data.frame() %>% 
  mutate(source="independent")
  r <- bind_rows(x_mcmc,
                 x_inde) %>% 
    mutate(source=as.factor(source))
  rand_samples <- sample(1:nrow(r), 0.7 * nrow(r))
  training_data <- r[rand_samples, ]
  testing_data <- r[-rand_samples, ]

  gbmFit1 <- train(source ~ ., data = training_data, 
                     method = "gbm",
                     trControl = trainControl(method = 'none'), 
                     tuneGrid = caretGrid, verbose=FALSE)

  plda <- predict(object=gbmFit1, newdata=testing_data)
  a_accuracy <- 
    tibble(predicted=plda, actual=testing_data$source) %>%
    mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
    summarise(mean(correct)) %>% 
    pull()
  return(a_accuracy)
}
library(furrr)
future::plan(multiprocess)
laccuracy <- future_map_dbl(seq(1, 1000, 1), function(i) f_distribution(x_mcmc))
hist(laccuracy, 50)
```

Could do same test with simply empirical distributions, also, by bootstrapping

## 10,000 iterations
Nominal model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_nom.stan', seed = 7878,
            refresh = 0, iter = 50000)
saveRDS(fit, "../output/fit_nominal_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_nom <- r_star_dist(full_data)
```

Alternative model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_alt_1.stan', seed = 7878,
            refresh = 0, iter = 50000)
saveRDS(fit, "../output/fit_alt_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_alt <- r_star_dist(full_data)
```

Plot
```{r}
b_df <- tibble(nominal=r_star_nom,
               alt=r_star_alt)

g1 <-
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        legend.position = "none") +
  ggtitle("A.")
g1

g2 <-
  b_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black")) +
  ggtitle("B.")
g2

pdf("../output/cauchy.pdf", width = 12, height = 6)
multiplot(g1, g2, cols = 2)
dev.off()
```

>>>>

r_star_nom <- r_star_dist(x)

mon <- monitor(fit_nom)
a_rhat_df <- tibble(nominal=mon$Rhat[1:50])
```

@@

r_star_alt <- r_star_dist(full_data)
mon <- monitor(fit_alt1)
a_rhat_df <-a_rhat_df %>% 
  mutate(alt=mon$Rhat[101:150],
         source="r_hat")
```

Plot split-R* values for each parameterisation
```{r}
a_df <- tibble(nominal=r_star_nom,
               alt=r_star_alt,
               source="r_star") %>% 
  bind_rows(a_rhat_df)
g <-
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        legend.position = "none") +
  facet_wrap(~source, scales = "free")
g
```

## 10,000 iterations
Nominal model with 50000 iterations and thinning by 5
```{r}
fit <- stan(file = 'cauchy_nom.stan', seed = 7878,
            refresh = 0, iter = 50000, thin=5)
saveRDS(fit, "../output/fit_nominal_50000.rds")
# fit <- readRDS("../output/fit_nominal_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_nom <- r_star_dist(full_data)
mon_nom <- monitor(fit)
```

Alternative model with 50000 iterations and thinning by 5
```{r}
fit <- stan(file = 'cauchy_alt_1.stan', seed = 7878,
            refresh = 0, iter = 60000, thin=3)
saveRDS(fit, "../output/fit_alt_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_alt <- r_star_dist(full_data)
mon_alt <- monitor(fit)
```

Plot
```{r}
b_df <- tibble(nominal=r_star_nom,
               alt=r_star_alt,
               source="r_star") %>% 
  bind_rows(tibble(nominal=mon_nom$Rhat[1:50],
                alt=mon_alt$Rhat[101:150],
                source="r_hat")) %>% 
  mutate(iter=10000)

c_df <- a_df %>% mutate(iter=1000) %>% 
  bind_rows(b_df) %>% 
  mutate(source=if_else(source=="r_hat", "split-Rhat", source)) %>% 
  mutate(source=if_else(source=="r_star", "R*", source))

my_lims <- function(x) { if (max(x) < 1.2) c(0.985, 1.03) else c(0.9, 3)}

g <- c_df %>% 
  melt(id.vars=c("source", "iter")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  facet_wrap(~source+iter, scales="free") +
  scale_x_continuous(limits=my_lims) +
  theme(strip.text = element_text(size=14, colour="black"),
        axis.text = element_text(size=12, colour="black"),
        axis.title = element_text(size=14, colour="black"),
        legend.text = element_text(size=14, colour="black"),
        legend.title = element_text(size=14, colour="black")) +
  xlab("Value")

ggsave("../output/cauchy.pdf", g, width = 12, height = 8)
```
